---
title: "HW 4_IE6600_Sec03_Group13"
author: "Vincenzo Coppola, Ashish Sharma, Akshay Dwivedi"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE}
require("knitr")
opts_knit$set(root.dir = "C:/Users/BigV24/Desktop/IE6600/Projects/MBTA Performance/data")
```

```{r Libraries, include=FALSE}
library(tidyverse)
library(dplyr)
library(lubridate)
library(ggplot2)
library(corrplot)
library(treemapify)
library(RColorBrewer)
library(ggalluvial)
library(ggfittext)
library(ggrepel)
library(scales)
```

```{r Data, include=FALSE}
df_satisfaction <- read.csv('MBTA_Service_Customer_Satisfaction.csv', na.strings="")
df_gatedentry <- read.csv('MBTA_Gated_Station_Entries.csv', na.strings="")
df_fueleff <- read.csv('Fuel_and_Energy.csv', na.strings="")
```

```{r Task 2, include=TRUE}
df_sat_split <- df_satisfaction %>% 
  separate(誰..survey_date, c('date','time'), sep = " ")
df_sat_split
monthly_only <- month(ymd(df_sat_split$date))
yearly_only <- year(ymd(df_sat_split$date))
df_satisfaction$MONTH <- monthly_only
df_satisfaction$YEAR <- yearly_only
st='%'
df_grouped <- df_satisfaction %>% 
  group_by(question_description,YEAR) %>% 
  summarise(Extremely_dissatisfied = mean(response_1_percent), Very_dissatisfied = mean(response_2_percent), Somewhat_dissatisfied = mean(response_3_percent), Neutral = mean(response_4_percent), Somewhat_satisfied = mean(response_5_percent), Very_satisfied = mean(response_6_percent), Extremely_satisfied= mean(response_7_percent),.groups='drop') %>% pivot_longer(c(Extremely_dissatisfied:Extremely_satisfied),names_to = "responses",values_to = 'percentages')

df_grouped$percentages <- label_percent()(df_grouped$percentages)


df_grouped
write.csv(df_grouped,"df_sat_flourish.csv", row.names = FALSE)
```

### Output 2
In the second task, corrplot was used to visualize the correlation between all delay types. By transforming the selected columns of the airline data into a matrix and ordering the matrix corrplot.mixed() could be used to visualize the correlation with two different methods - area circles, and numeric values. The resulting correlation matrix shown in the output is the visualization from relating the delay types to each other with a 0-1 range for numbers and circle size where an increase is increase in correlation.

```{r Task 3, include = TRUE}
colnames(df_gatedentry)
df_entry_split <- df_gatedentry %>% 
  separate(誰..service_date, c('date','time'), sep = " ")
head(df_entry_split)
monthly_only2 <- month(ymd(df_entry_split$date))
yearly_only2 <- year(ymd(df_entry_split$date))
df_entry_split$MONTH <- monthly_only2
df_entry_split$YEAR <- yearly_only2
df_groupedentry <- df_entry_split %>% 
  group_by(station_name,route_or_line,YEAR) %>% 
  summarise(Total_entries = sum(gated_entries),.groups='drop')
head(df_groupedentry)

df_groupedentry_race <- df_entry_split %>% 
  group_by(route_or_line,time_period) %>% 
  summarise(Total_entries = sum(gated_entries),.groups='drop') %>% 
  group_by(route_or_line) %>% 
  mutate(cumulative_entries = cumsum(Total_entries)) %>% 
  select(-Total_entries) %>% 
  pivot_wider(names_from = time_period, values_from = cumulative_entries)
  

head(df_groupedentry_race)

write.csv(df_groupedentry,"df_gatedentry_floruish.csv", row.names = FALSE)
write.csv(df_groupedentry_race,"df_barchart_floruish.csv", row.names = FALSE)


df_dips <- df_entry_split %>% 
  group_by(route_or_line,time_period) %>% 
  summarise(Total_entries = sum(gated_entries),.groups='drop') %>% 
  pivot_wider(names_from = route_or_line, values_fill = Total_entries,)
(df_dips)
write.csv(df_dips,"df_dips_floruish.csv", row.names = FALSE)

df_usage_fuel <- df_entry_split %>% 
  filter(YEAR == '2016'|YEAR == '2017'|YEAR == '2018'|YEAR == '2019') %>% 
  group_by(YEAR) %>% 
  summarise(Total_entries = sum(gated_entries),.groups='drop')

df_usage_fuel_combo <- df_fueleff_emission %>% 
  group_by(誰..Year) %>% 
  summarise(Total_vol = sum(Volume,na.rm = TRUE),Total_Miles = sum(Miles_Traveled,na.rm = TRUE), avg_eff=mean(Efficiency,na.rm = TRUE),Total_emissions = sum(Emissions,na.rm = TRUE),.groups='drop')

df_withgasonly <- df_usage_fuel_combo %>% 
  mutate(Emissions_if_gas = 19.37*Total_vol)

df_withgasonly

df_finalfuel1 <- bind_cols(df_withgasonly,df_usage_fuel)
df_finalfuel <- df_finalfuel1 %>% 
  mutate(drivers = Total_entries/(4.1*365*2*115)) %>% 
  mutate(milesbydrivers = (drivers*Total_Miles/4)) %>% 
  mutate(emissions_if_drivers = milesbydrivers*.9061)

df_finalfuel
write.csv(df_finalfuel,"df_fueldiff_floruish.csv", row.names = FALSE)
```


```{r Task 3, include = TRUE}
colnames(df_fueleff)
df_diesel <- df_fueleff %>% 
  filter(Fuel_Source=='Diesel'|Fuel_Source=='Bio-Diesel') %>% 
  mutate(Emissions = 22.46 * Volume)

df_diesel

df_gas <- df_fueleff %>% 
  filter(Fuel_Source=='Gasoline'|Fuel_Source=='Liquefied Petroleum Gas'|Fuel_Source=='Other Fuel') %>% 
  mutate(Emissions = 19.37 * Volume)

df_gas

df_electric <- df_fueleff %>% 
  filter(Fuel_Source=='Electric Propulsion'|Fuel_Source=='Electric Battery') %>% 
  mutate(Emissions = 0.85 * Volume)

df_electric

df_naturalgas <- df_fueleff %>% 
  filter(Fuel_Source=='Compressed Natural Gas') %>% 
  mutate(Emissions = 16.17 * Volume)

df_naturalgas

df_fueleff_emission1 <- bind_rows(df_naturalgas,df_electric)
df_fueleff_emission2 <- bind_rows(df_fueleff_emission1,df_gas)
df_fueleff_emission <- bind_rows(df_fueleff_emission2,df_diesel)
dim(df_fueleff_emission)

df_grouped_efficiency <- df_fueleff_emission %>% 
  group_by(誰..Year,Fuel_Source) %>% 
  summarise(Total_vol = sum(Volume,na.rm = TRUE),Total_Miles = sum(Miles_Traveled,na.rm = TRUE), avg_eff=mean(Efficiency,na.rm = TRUE),Total_emissions = sum(Emissions,na.rm = TRUE),.groups='drop')

df_grouped_efficiency

write.csv(df_grouped_efficiency,"df_efficiency_floruish.csv", row.names = FALSE)

df_grouped_noyear <- df_fueleff_emission %>% 
  group_by(Fuel_Source) %>% 
  summarise(Total_vol = sum(Volume,na.rm = TRUE),Total_Miles = sum(Miles_Traveled,na.rm = TRUE), avg_eff=mean(Efficiency,na.rm = TRUE),Total_emissions = sum(Emissions,na.rm = TRUE),.groups='drop')
df_grouped_noyear


write.csv(df_grouped_noyear,"df_efficiency_datawrap.csv", row.names = FALSE)
```
#Sources
1. https://datatofish.com/export-dataframe-to-csv-in-r/
2. https://data.boston.gov/dataset/crime-incident-reports-august-2015-to-date-source-new-system
3. https://www.google.com/covid19/mobility/


